---
title: CPRD Mouth Ulcer Case Control Study
author: "Pieta"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{CPRD Mouth Ulcer Case Control Study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Motivation

Use the Mouth Ulcers Case-Control Data to develop R package to load and clean CPRD Aurum data and look
for codes "enriched" in patients with a particular outcome prior to their first outcome code.


## Setup

### Paths

```{r}
require(plib)
require(conflicted)
require(duckdb)
require(cprdaurumtools)
require(tidyverse)
homeDir <- Sys.getenv("HOME")
projRoot <- file.path(homeDir,"Projects")
projdir <- file.path(projRoot,"mung")
dir.create(projdir,showW=F)
datadir <- file.path(projdir,".data","case_control")
cprddir <- file.path(projRoot,"refdata","aurum","202309_Lookups_CPRDAurum")
dbfile <- file.path(datadir,"case_control.duckdb")
.disp <- T
options(warn=F)
```

### Load Files

```{r}
load_patients(pdir=datadir,dbf=dbfile,ow=F)
load_cprdfiles(pddir=cprddir,dbf=dbfile,ow=F)
load_cons(pddir=datadir,dbf=dbfile,ow=F)
load_obs(pddir=datadir,dbf=dbfile,ow=F)
load_drugissues(pddir=datadir,dbf=dbfile,ow=F)
load_practice(pddir=datadir,dbf=dbfile,ow=F)
load_staff(pddir=datadir,dbf=dbfile,ow=F)
load_probs(pddir=datadir,dbf=dbfile,ow=F)
load_referrals(pddir=datadir,dbf=dbfile,ow=F)
```

## Load Medcodes 

### Cases

```{r, results="asis", comment=NA}
cohortcodes <- read_tsv(file.path(projdir,"refs","bis.txt"),
                      col_type=cols(.default=col_character()))
dbi <- duckdb::dbConnect(duckdb::duckdb(),dbfile)
if(F){
  dbListTables(dbi)
  ppiesymptomcodes <- dbGetQuery(dbi,
    str <- c("SELECT e.medcodeid,e.snomedctconceptid,e.emiscodecategoryid, e.term
             FROM emismedicaldictionary AS e 
             INNER JOIN ppiesymptomslist AS p
             ON p.medcodeid=e.medcodeid")) %>% tibble() %>%
  group_by(medcodeid,snomedctconceptid,emiscodecategoryid) %>% 
  summarise(term=first(term),.groups="drop") %>%
  dplyr::filter(as.numeric(emiscodecategoryid)>=11 & as.numeric(emiscodecategoryid)<33) %>%
  arrange(emiscodecategoryid)
  write_csv(file=file.path(projdir,"refs","ppiesymptoms.csv"))
}
dbWriteTable(dbi,"cohortcodes",cohortcodes,overwrite=T)
duckdb::dbDisconnect(dbi)
```

### Symptoms

```{r}
symptomcodes<- read_csv(file.path(projdir,"refs","ulcer.csv"),
                      col_type=cols(.default=col_character()))
names(symptomcodes) <- tolower(names(symptomcodes))

dbi <- duckdb::dbConnect(duckdb::duckdb(),dbfile)
dbWriteTable(dbi,"symptomcodes",symptomcodes,overwrite=T)
duckdb::dbDisconnect(dbi)
```

## Assign Patients To Cohorts

Patients are cases or controls but there are several types of case and I have realised that the controls
will be depleted for patients of the type in the other cases if I am not careful and while with SLE or
Behcet's that might not make a big difference that might make a difference if depleted for IBD as there
are many more IBD and in comparison to the SLE and Behect's numbers this might be significant

```{r, results="asis", comment=NA}
ddb <- duckdb::dbConnect(duckdb(),dbfile)
if('.id'%in%dbListFields(ddb,"patients")){
  ssql <- str_c("ALTER TABLE patients RENAME COLUMN '.id' TO cohort")
  dbExecute(ddb,ssql)
}
duckdb::dbDisconnect(ddb,shutdown=T)
```

Get the cohorts numbers

```{r}
cohorts <- cprdaurumtools::get_cohort(dbfile) %>% select(-cohort)
indexdates <- get_cohort(dbfile,indexfield="indexdate",vfill=NA)
indexages <- get_cohort(dbfile,indexfield="indexage",vfill=NA)
indexagesum <- indexages %>% pivot_longer(-c(patid,cohort),names_to="codeclass",values_to="indexage") %>%
  dplyr::mutate(hit=ifelse(is.na(indexage),0,1)) %>%
  group_by(cohort,codeclass) %>%
  dplyr::summarise(cases=sum(hit),meanage=mean(indexage,na.rm=T),.groups="drop")
idxagesum <- indexagesum %>% select(codeclass,meanage)
indexagesum %>% display_data(disp=.disp,buttons=T,)
```

### Spiking the Controls

There are a total of about 22 million eligible patients born in the range of dates the cases and controls
are sampled from. Because we have sampled about 100 thousand patients from this range that means we would
probably only expect 1 SLE case in the controls for Behcet's and IBD, no Behcet's in the controls for SLE
and IBD but maybe 50 IBD cases in the controls for SLE and Behcet's, but because the controls were
depleted for all three they are not in just the control sample however to include all the SLE cases in
the controls for either IBD or Behcet's case vastly over enriches the controls for SLE. This is
especially true with IBD, to include them is to vastly over enrich the controls for IBD cases but to
exclude them is also possibly depleting the controls too much.

There for we need to add about 50 random IBD cases to the controls for SLE and Behcets

```{r}
# select random IBD cases for controls
ibdconts <- cohorts %>% dplyr::filter(ibd>0 & sle<1 & behcets<1) %>% pull(patid) %>% sample(50)
pats <- get_patients(dbfile,fields=c("patid")) %>% left_join(cohorts, by=c("patid")) %>% tibble()
pats[is.na(pats)] <- 0
patscohort <- pats %>%  
  dplyr::mutate(
         ibdcohort=ifelse(ibd>0,"case",ifelse(sle<1 & behcets <1, "cont","drop")),
         slecohort=ifelse(sle>0,"case",
                          ifelse((ibd<1 & behcets <1),"cont",
                                ifelse(patid %in% ibdconts,"cont","drop"))),
         behcetscohort=ifelse(behcets>0,"case",
                          ifelse((ibd<1 & sle <1),"cont",
                                ifelse(patid %in% ibdconts,"cont","drop")))) %>%
  select(patid,ibdcohort,slecohort,behcetscohort) %>% 
  dplyr::rename(ibd=ibdcohort,sle=slecohort,behcets=behcetscohort)
  
patscohort %>% pivot_longer(-patid,names_to="cohort",values_to="cc") %>%
  group_by(cohort,cc) %>% dplyr::summarise(n=n(),.groups="drop") %>% 
  pivot_wider(names_from="cc",values_from="n",values_fill=0) %>% 
  dplyr::mutate(total=case+cont) %>% display_data(disp=.disp,buttons=T,)
```

### Generate Pseudo Index Ages

```{r}
indexageslong <- indexages %>% select(-cohort) %>% 
  pivot_longer(-patid,names_to="cohort",values_to="indexage") %>% na.omit()
indexagetab <- patscohort %>% pivot_longer(-patid,names_to='cohort',values_to='status') %>%
  dplyr::filter(status!='drop') %>% inner_join(idxagesum, by=c("cohort"="codeclass")) %>%
  left_join(indexageslong,by=c('patid','cohort')) %>%
  dplyr::mutate(indexage=ifelse(is.na(indexage),meanage,indexage)) %>%
  select(patid,cohort,status,indexage)
ddb <- duckdb::dbConnect(duckdb(),dbfile)
dbListTables(ddb)
if('cohorts'%in%dbListTables(ddb)){
  dbWriteTable(ddb,'cohorts',indexagetab,over=T)
}
duckdb::dbDisconnect(ddb,shutdown=T)
```

## Results

### Individual Ulcer Code List 

```{r}
coh <- 'sle'
res <- run_test(dbf,cohort=coh,symptomtab="symptomcodes",cortype="bon")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
res <- run_test(dbf,cohort=coh,symptomtab="symptomcodes",groupat="snomedctconceptid")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
```

```{r}
coh <- 'ibd'
res <- run_test(dbf,cohort=coh,symptomtab="symptomcodes")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
res <- run_test(dbf,cohort=coh,symptomtab="symptomcodes",groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
```

```{r}
coh <- 'behcets'
res <- run_test(dbf,cohort=coh,symptomtab="symptomcodes") 
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
res <- run_test(dbf,cohort=coh,symptomtab="symptomcodes",groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
```


### PPIE Symptoms

Look at the expanded codes that came from the PPIE

```{r}
symptomtable <- "ppiesymptomslist"
coh <- 'sle'
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable,cortype="bon")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
```

```{r}
coh <- 'ibd'
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable)
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
```

```{r}
coh <- 'behcets'
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable) 
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
```

### Expanded Symptoms

Look at the expanded codes that came from the PPIE

```{r}
symptomtable <- "expandedsymptomlist"
coh <- 'sle'
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable,cortype="bon")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
```

```{r}
coh <- 'ibd'
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable)
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
```

```{r}
coh <- 'behcets'
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable) 
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
res <- run_test(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,case,cont,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
```

## Correct Index Age 

After discussing with Liv and Clare the issue of the pap vaccination be uplifted in IBD it is clear that
this is potentially an issue with the fixed index age need to generate a distribution of index dates not
a fixed mean

```{r}
res <- gen_indexdist(dbf=dbfile)
indexages <- get_indexdist(dbf=dbfile)
```

### SLE

```{r, results='hide'}
coh <- "sle"
indexages %>% dplyr::filter(cohort=="cases",codetype==coh) %>% pull(iage) %>%
  ecdf() %>% plot(xlab="age",ylab="CDF",main=paste0("CDF of ",coh," cases"))
indexages %>% dplyr::filter(cohort=="conts",codetype==coh) %>% pull(iage) %>%
  ecdf() %>% plot(xlab="age",ylab="CDF",main=paste0("CDF of ",coh," controls"))
```

### IBD

```{r, results='hide'}
coh <- "ibd"
indexages %>% dplyr::filter(cohort=="cases",codetype==coh) %>% pull(iage) %>%
  ecdf() %>% plot(xlab="age",ylab="CDF",main=paste0("CDF of ",coh," cases"))
indexages %>% dplyr::filter(cohort=="conts",codetype==coh) %>% pull(iage) %>%
  ecdf() %>% plot(xlab="age",ylab="CDF",main=paste0("CDF of ",coh," controls"))
```

### Behcet's

```{r, results='hide'}
coh <- "behcets"
indexages %>% dplyr::filter(cohort=="cases",codetype==coh) %>% pull(iage) %>%
  ecdf() %>% plot(xlab="age",ylab="CDF",main=paste0("CDF of ",coh," cases"))
indexages %>% dplyr::filter(cohort=="conts",codetype==coh) %>% pull(iage) %>%
  ecdf() %>% plot(xlab="age",ylab="CDF",main=paste0("CDF of ",coh," controls"))
```

But now with this new fullindexage table I need to respike the controls. Probably remove all controls
with SLE or Behcet's and add back 50 or so IBD controls for SLE and Behcet's

```{r}
spike_cohorts <- function(dbf,tabname="studycohorts"){
  cohorts <- get_indexdist(dbf=dbfile) %>% select(patid,cohort,codetype) %>% 
  pivot_wider(names_from="codetype",values_from="cohort") %>%
  mutate(ibd=ifelse(ibd=="cases",ibd,ifelse(sle=="cases" | behcets=="cases","drop",ibd)),
         sle=ifelse(sle=="cases",sle,ifelse(behcets=="cases" | ibd=="cases","drop",sle)),
         behcets=ifelse(behcets=="cases",behcets,ifelse(sle=="cases" | ibd=="cases","drop",behcets)))

  ibdspikes <- cohorts %>% dplyr::filter(ibd=="cases"& sle=="drop" & behcets=="drop") %>% pull(patid) %>%
  sample(50)

  cohorts <- cohorts %>% mutate(sle=ifelse(patid %in% ibdspikes,"conts",sle),
                     behcets=ifelse(patid %in% ibdspikes,"conts",behcets)) %>%
                     pivot_longer(-patid,names_to="study",values_to="cohort") %>%
                     dplyr::filter(cohort!="drop")
  ddb <- duckdb::dbConnect(duckdb::duckdb(),dbfile)
  dbWriteTable(ddb,tabname,cohorts,overwrite=T)
  dbDisconnect(ddb,shutdown=T)
}
ddb <- duckdb::dbConnect(duckdb::duckdb(),dbfile)
tabs <- dbListTables(ddb)
dbDisconnect(ddb,shutdown=T)
if(!"studycohorts"%in%tabs){
  spike_cohorts(dbfile)
}
```

### Individual Ulcer Code List 

```{r}
coh <- 'sle'
res <- run_test2(dbf,cohort=coh,symptomtab="symptomcodes",cortype="bon")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
res <- run_test2(dbf,cohort=coh,symptomtab="symptomcodes",groupat="snomedctconceptid")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
```

```{r}
coh <- 'ibd'
res <- run_test2(dbf,cohort=coh,symptomtab="symptomcodes")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
res <- run_test2(dbf,cohort=coh,symptomtab="symptomcodes",groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
```

```{r}
coh <- 'behcets'
res <- run_test2(dbf,cohort=coh,symptomtab="symptomcodes") 
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
res <- run_test2(dbf,cohort=coh,symptomtab="symptomcodes",groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
```


### PPIE Symptoms

Look at the expanded codes that came from the PPIE

```{r}
symptomtable <- "ppiesymptomslist"
symptomtab <- "ppiesymptomslist"
coh <- 'sle'
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable,cortype="bon")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
```

```{r}
coh <- 'ibd'
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable)
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
```

```{r}
coh <- 'behcets'
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable) 
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
```

### Expanded Symptoms

Look at the expanded codes that came from the PPIE

```{r}
symptomtable <- "expandedsymptomlist"
coh <- 'sle'
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable,cortype="bon")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="SLE patients")
```

```{r}
coh <- 'ibd'
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable)
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>%
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="IBD patients")
```

```{r}
coh <- 'behcets'
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable) 
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
res <- run_test2(dbf,cohort=coh,symptomtab=symptomtable,groupat="snomedctconceptid")
if(nrow(res)>0) res %>% 
  select(term,snomedctconceptid,cases,conts,ratio,padj) %>%
  display_data(disp=.disp,buttons=T,caption="Behcet's patients")
```

# Appendix

## Codes

### SLE

```{r}
cohortcodes %>% dplyr::filter(codetype=='sle') %>% select(medcodeid,term) %>% 
  dplyr::arrange(term) %>% kableExtra::kbl()
```

### Behcet's

```{r}
cohortcodes %>% dplyr::filter(codetype=='behcets') %>% select(medcodeid,term) %>%
  dplyr::arrange(term) %>% kableExtra::kbl()
```

### IBD

```{r}
cohortcodes %>% dplyr::filter(codetype=='ibd') %>% select(medcodeid,term) %>% 
  dplyr::arrange(term) %>% kableExtra::kbl()
```

### Ulcers

```{r}
symptomcodes %>% select(medcodeid,term) %>% 
  dplyr::arrange(term) %>% kableExtra::kbl()
```

# Other approaches

## Counts of Codes

```{r}
devtools::reload(pkg="/home/pietas/GitLab/cprdaurumtools")
help(sum_observations)
sum_obs <- sum_observations(dbfile,cases="sle")
sle_pats <- get_cohort(dbfile) %>% dplyr::filter(sle>0) %>% pull(patid)
sum_bycode <- sum_obs %>% mutate(cc=ifelse(patid%in%sle_pats,"sle","cont")) %>% group_by(cc,medcodeid) %>%
  summarise(n=n(),minfreq=min(freq),maxfreq=max(freq),meanfreq=mean(freq),.groups="drop")
sum_cc <- sum_obs %>% select(patid) %>% unique() %>%
  mutate(cc=ifelse(patid%in%sle_pats,"sle","cont")) %>%
  group_by(cc) %>% summarise(n=n(),.groups="drop")
meanfreq <- sum_bycode %>% 
  pivot_wider(id_cols=medcodeid,names_from="cc",values_from="meanfreq",values_fill=0) %>%
  mutate(test=sle/(cont+sle))
cohort_count <- sum_bycode %>% 
  pivot_wider(id_cols=medcodeid,names_from="cc",values_from="n",values_fill=0) %>%
  mutate(test_n=sle/(cont+sle)) %>% rename(cont_n=cont,sle_n=sle)
hightest <- meanfreq %>% arrange(desc(sle),desc(test)) %>% dplyr::filter(test>0.5) %>%
  inner_join(get_medcodes(dbf), by="medcodeid") %>% 
  select(medcodeid,term,sle,cont,test,snomedctconceptid) %>% 
  inner_join(cohort_count,by="medcodeid")
hightest %>% dplyr::filter(sle_n>=50) %>% display_data(disp=.disp,buttons=T)

get_medcodes(dbfile) %>% dplyr::filter(snomedctconceptid=="498521000006103")
```

This is interesting I think as it shows there are some differences. But I think I need to find some means
of looking for discriminants

Possibly age at first diagnosis

Let's first look at it the other way what comes up if we look at other

```{r}
lowtest <- meanfreq %>% arrange(desc(cont),test) %>% dplyr::filter(test<0.5) %>%
  inner_join(get_medcodes(dbf), by="medcodeid") %>% 
  select(medcodeid,term,sle,cont,test,snomedctconceptid) %>% 
  inner_join(cohort_count,by="medcodeid")
lowtest %>% dplyr::filter(sle_n<2) %>% display_data(disp=.disp,buttons=T)
```

This is not looking very revealing it appears that the number of times things are coded is about the
same.

Perhaps I should be looking at consultations not observations

```{r}

```




```{r echo=F,eval=F}
homeDir <- Sys.getenv("HOME")
inputpath <- file.path(homeDir,"GitLab","cprdaurumtools","vignettes")
inputFile <- file.path(inputpath,"mouth_ulcer_case_control.Rmd")
noteDir <- file.path(homeDir,"Notes","uol","mung")
system2("firefox",
        args=rmarkdown::render(inputFile,encoding=encoding,output_dir=noteDir))
```
