---
title: "cprdaurumtools Package development"
author: "Pieta"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Motivation

I currently have several CPRD project using Aurum, I will also have a GOLD project soon. There are some
common steps that need to be achieved with these projects. Mainly the ability to link HES data and to be
able to clean values in the observations tables.

The idea is that if I can import the data into an sqlite database it will permit the use of SQL to join
and extract data for the process of linkage and cleaning. 

It is possible that a second sqlite database will be created for the analysis data but the pre-processing
can be done partly in SQL. These project are all going to be too large to do the type of queries I want
in memory.

## Setup

### Paths

```{r}
require(cprdaurumtools)
require(tidyverse)
homeDir <- Sys.getenv("HOME")
projRoot <- file.path(homeDir,"Projects")
pdir <- file.path(projRoot,"cprdaurumtools")
dir.create(pdir,showW=F)
batch <- 1 
ddir <- file.path(projRoot,"t2dd",".data","aurum_may",paste0("batch",batch))
rdir <- file.path(projRoot,"cprd","refs","202205_Lookups_CPRDAurum")
ldir <- file.path(projRoot,"t2dd",".data","linked_data")
hdir <- file.path(ldir,"hes")
odir <- file.path(ldir,"ons")
idir <- file.path(ldir,"imd")
dbif <- file.path(pdir,".data","aurumdata_prep.sqlite")
dir(ddir)
dir(rdir)
dir(hdir) 
dir(idir)
dir(odir)
```

### Load Patients

```{r}
pddir <- file.path(ddir,"Patient")
load_patients(pddir=pddir,dbf=dbif,ow=F)
str_sql <- "SELECT patid FROM patients;"
dbi <- RSQLite::dbConnect(RSQLite::SQLite(),dbif)
patids <- RSQLite::dbGetQuery(dbi,str_sql) %>% as_tibble() %>% pull(patid)
RSQLite::dbDisconnect(dbi)
```

## Load Observations

```{r}
pddir <- file.path(ddir,"Observation")
load_obs(pddir=pddir,dbf=dbif,ow=F)
```

## Load CPRD references

```{r}
load_cprdfiles(pddir=rdir,dbf=dbif,ow=F)
```

## Load Drug Issues

```{r}
pddir <- file.path(ddir,"DrugIssue")
load_drugissues(pddir=pddir,dbf=dbif,ow=F)
```

## Load HES data

```{r}
load_hesfiles(pddir=hdir,dbf=dbif,tad="21_001631.*",pats=patids,ow=F)
```

## Load IMD data

```{r}
load_imdfiles(pddir=idir,dbf=dbif,tad="21_001631.*",pats=patids,ow=F)
```

## Load ONS data

```{r}
load_onsfiles(pddir=odir,dbf=dbif,tad="21_001631.*",pats=patids,ow=F)
```

## Pre-processing

At this point we have loaded a lot of data into the pre-processing database the next step is to transfer
processed records to the analysis database

### Complete Cases

Possibly we need to join the patient table with the HES patient table and their ONS and IMD data

```{r}
dbanal <- file.path(pdir,".data","aurumdata_anal.sqlite")
dbprep <- file.path(pdir,".data","aurumdata_prep.sqlite")
proc_patients(dbpre=dbprep,dbpost=dbanal)
```



